R for Clinical Study Reports and Submission

### 2021-09-19 tlf-disposition.Rmd
revert analysis code

The first step is to read in the relevant datasets into R.
For a disposition table, all the required information is saved in a Subject-level Analysis Dataset (ADSL).
The dataset is provided in SAS7bdat format, which is the current standard in current clinical trial reporting and analysis.The `haven` package is able to read in the dataset, while maintaining its attributes (eg. variable labels).

@@

-   USUBJID: Unique subject identifier
-   TRT01P: Planed treatment
-   TRT01PN: Planned treatment numeric encoding
-   DISCONFL: Discontinued from study flag
-   DCREASCD: Discontinued from Study reason coded

@@

```{r}
n_rand <- adsl %>%
  group_by(TRT01P) %>%
  summarize(n = n()) %>%
  pivot_wider(
    names_from = TRT01P,
    names_prefix = "",
    values_from = n
  ) %>%
  mutate(row = "Participants Randomized")

@@

```{r}
n_disc <- adsl %>%
  group_by(TRT01P) %>%
  summarize(
    n = sum(DISCONFL == "Y"),
    pct = formatC(n / n() * 100,
      digits = 1, format = "f", width = 5
    )
  ) %>%
  mutate(npct = paste0(n, " (", pct, "%)")) %>%
  select(-c(n,pct)) %>%
  pivot_wider(
    names_from = TRT01P,
    values_from = npct,
    values_fill = list(npct = "0")
  ) %>%
  mutate(row = "Discontinued")

@@

```{r}
n_reason <- adsl %>%
  group_by(TRT01P) %>%
  mutate(n_total = n()) %>%
  group_by(TRT01P, DCREASCD) %>%
  summarize(
    n = n(),
    pct = formatC(n / unique(n_total) * 100,
      digits = 1, format = "f", width = 5
    )
  ) %>%
  mutate(npct = paste0(n, " (", pct, "%)")) %>%
  pivot_wider(
    id_cols = DCREASCD,
    names_from = TRT01P,
    values_from = npct,
    values_fill = list(npct = "0")
  ) %>%
  rename(row = DCREASCD)

@@

```{r}
tbl_disp <- rbind(n_rand, n_complete, n_disc, n_reason)

@@

  # Table title
  rtf_title("Disposition of Participants") %>%

>>>>

The first step is to read in the relevant datasets into R.
For a disposition table, all the required information is saved in a Subject-level Analysis Dataset (ADSL).
The dataset is provided in `sas7bdat` format that is a SAS data format currently used in many clinical trial analysis and reporting.
The `haven` package is able to read in the dataset, while maintaining its attributes (e.g. variable labels).

@@

-   USUBJID: Unique subject identifier
-   TRT01P: Planed treatment
-   TRT01PN: Planned treatment numeric encoding
-   DISCONFL: Discontinued from study flag
-   DCREASCD: Discontinued from study reason coded

@@

```{r}
n_rand <- adsl %>%
  group_by(TRT01PN) %>%
  summarize(n = n()) %>%
  pivot_wider(
    names_from = TRT01PN,
    names_prefix = "n_",
    values_from = n
  ) %>%
  mutate(row = "Participants in population")

@@

```{r}
n_disc <- adsl %>%
  group_by(TRT01PN) %>%
  summarize(
    n = sum(DISCONFL == "Y"),
    pct = formatC(n / n() * 100,
      digits = 1, format = "f", width = 5
    )
  ) %>%
  pivot_wider(
    names_from = TRT01PN,
    values_from = c(n, pct)
  ) %>%
  mutate(row = "Discontinued")

@@

```{r}
n_reason <- adsl %>%
  group_by(TRT01PN) %>%
  mutate(n_total = n()) %>%
  group_by(TRT01PN, DCREASCD) %>%
  summarize(
    n = n(),
    pct = formatC(n / unique(n_total) * 100,
      digits = 1, format = "f", width = 5
    )
  ) %>%
  pivot_wider(
    id_cols = DCREASCD,
    names_from = TRT01PN,
    values_from = c(n, pct),
    values_fill = list(n = 0, pct = "  0.0")
  ) %>%
  rename(row = DCREASCD)

@@

```{r}
tbl_disp <- bind_rows(n_rand, n_complete, n_disc, n_reason) %>%
  select(row, ends_with(c("_0", "_54", "_81")))

@@

  # Table title
  rtf_title("Disposition of Patients") %>%
